Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-08-07 Paper/assa.tex
..

\label{sub:gsa}
Consider a function: $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
While the active subspace framework described above does not make any assumptions
about statistical independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
the classical 
framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} 
assumes that the inputs
are statistically independent. While extensions to the cases 
of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
 we limit the discussion in this section to the
case of random inputs that are statistically independent and are 
either uniformly distributed or
distributed according to a Boltzmann probability distribution.
Note that a measure $\mu$
on $\R$ is referred to as a Boltzmann measure if it is 
absolutely continuous with respect to the Lebesgue measure  
and admits a density  of the form $\pi(x) = C \exp\{-V(x)\}$,
where $V$ is a continuous function and $C$ a normalization 
constant~\cite{Lamboni:2013}.

>>>>

\label{sub:gsa}
Consider a function: $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
While the active subspace framework described above does not make any assumptions
about statistical independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
the classical 
framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} 
assumes that the inputs
are statistically independent. While extensions to the cases 
of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007,Hart:2017},
 we limit the discussion in this section to the
case of random inputs that are statistically independent and are 
either uniformly distributed or
distributed according to a Boltzmann probability distribution.
Note that a measure $\mu$
on $\R$ is referred to as a Boltzmann measure if it is 
absolutely continuous with respect to the Lebesgue measure  
and admits a density  of the form $\pi(x) = C \exp\{-V(x)\}$,
where $V$ is a continuous function and $C$ a normalization 
constant~\cite{Lamboni:2013}.

### 2018-08-07 Paper/method.tex
..

%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative $L_2$ 
norm of the difference between
the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvectors. In this case, since 
$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we next explore the applicability of a so-called gradient-free approach
as mentioned earlier.

>>>>

%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. Right: Relative $L_2$ 
norm of the difference between
the square of individual components of the eigenvector evaluated using $N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
The comparison is illustrated for the first five eigenvectors.} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
the higher-order eigenvectors. In this case, since 
$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
case. The iterative strategy therefore offers a significant potential for computational advantage. 
However, estimating the gradient of the model output imposes a significant computational burden.
Hence, we next explore the applicability of a so-called gradient-free approach
as mentioned earlier.
